Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-08-29 Manuscript.Rmd
PDF images problem fixed

**Meta-analysis**
We used data from metan-alyses that gathered peak brain coordinates of activity from studies of valuation (Bartra et al., 2010) (n = 27 studies) and DMN (Laird et al., 2009) (n = 77). These represent the surviving areas post-statistical thresholding from each study. For each peak volumetric coordinate, 10 mm sphere masks were generated and then projected to a standard cortical mesh (fs_average, 32,000 vertices) using FreeSurfer. The sphere sections that overlapped with the cortex (should I add the details of the overlap?) were mapped to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study.

@@

**fMRI Data**
In order to quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state fMRI data from the Human Connectome Project (HCP; Van Essen et al., 2013; www.humanconnectome.org) Q6 release (n = 80, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state sessions (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) were each comprised of 1200 TRs for a total scan time of 14 min 33 s, with left-right and left-right phase encoding implemented on each day. We used this scanning structure to estimate each subject's topology based on session, day, and overall data (4800 TRs).

@@

**Within-ROI Network Definition**
To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,564 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately; 2) the concatenated timeseries from each pair of daily sessions; 3) each subject's dataset; and 4) on each step of a sliding window analysis (see Evaluation for more details).

**Community Detection**

@@

**Partition Evaluation**
To quantify the agreement between partitions, we computed both the adjusted rand index and variation of information distance. The adjusted rand index denotes the proportion of vertices that coincide in affiliation across partitions, and compares this score to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016).

@@

## Results
## Meta-analysis

@@

```{r Meta-analysis ROIs, out.width="60%", out.height="40%", fig.align = "center"}
# include_graphics("./Images/ROIS.png")
```

## Partition Comparison: Full Datasets

@@

```{r Template partition, out.width="50%", out.height="30%", fig.align = "center"}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nSubj <- length(Summaries)

#include_graphics("./Images/template.png")
```

@@

## Partition Comparison: Day 1 vs Day 2

@@

## Partition Comparison: Session

@@

## Stability per subject

@@

## Correlation vs community detection in mPFC

@@

## Structural correspondence

>>>>

### *Meta-analysis*

We used data from metan-alyses that gathered peak brain coordinates of activity from studies of valuation (Bartra et al., 2010) (n = 27 studies) and DMN (Laird et al., 2009) (n = 77). These represent the surviving areas post-statistical thresholding from each study. For each peak volumetric coordinate, 10 mm sphere masks were generated and then projected to a standard cortical mesh (fs_average, 32,000 vertices) using FreeSurfer. The sphere sections that overlapped with the cortex (should I add the details of the overlap?) were mapped to an atlas of the human cortical surface (Glasser et al., 2016). This produced a list of standardized parcels that were reported on each study.

@@

### *fMRI Data*

In order to quantify the intrinsic connectivity of these ROIs in the brain, we acquired resting-state fMRI data from the Human Connectome Project (HCP; Van Essen et al., 2013; www.humanconnectome.org) Q6 release (n = 80, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state sessions (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) were each comprised of 1200 TRs for a total scan time of 14 min 33 s, with left-right and left-right phase encoding implemented on each day. We used this scanning structure to estimate each subject's topology based on session, day, and overall data (4800 TRs).

@@

### *Within-ROI Network Definition*

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,564 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we took the exponential of the correlations, Fisher transformed them (i.e. tanh) so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately; 2) the concatenated timeseries from each pair of daily sessions; 3) each subject's dataset; and 4) on each step of a sliding window analysis (see Evaluation for more details).

### *Community Detection*

@@

### *Partition Evaluation*

To quantify the agreement between partitions, we computed both the adjusted rand index and variation of information distance. The adjusted rand index denotes the proportion of vertices that coincide in affiliation across partitions, and compares this score to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016).

@@

\newpage

## Results

### *Meta-analysis*

@@

```{r Meta-analysis ROIs, out.width="60%", fig.align = "center"}
include_graphics("./Images/ROIS.png")
```

### *Partition Comparison: Full Datasets*

@@

```{r Template partition, out.width="50%", fig.align = "center"}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nSubj <- length(Summaries)

include_graphics("./Images/template.png")
```

@@

### *Partition Comparison: Day 1 vs Day 2*

@@

### *Partition Comparison: Session*

@@

### *Stability per subject*

@@

### *Correlation vs community detection in mPFC*

@@

### *Structural correspondence*
